\section{Conclusion}

This report described the full pipeline from a relational \ac{imdb} source to a deployed and queryable knowledge graph. Milestone~1 produced a 14-relation schema reduced from a conceptual \ac{erd} and populated from a numVotes-ranked top-$N$ slice of the official \ac{imdb} non-commercial dataset. Milestone~2 produced an \ac{owl}~2 \ac{dl} ontology centred on \texttt{CreativeWork} as the abstraction of all titles, with a reified \texttt{Participation} class capturing per-credit context. Milestone~3 connected the two layers via an \ac{r2rml} mapping that produced a 1{,}611{,}676-triple knowledge graph. Milestone~4 deployed the result in Apache Fuseki, evaluated a battery of \ac{sparql} queries covering subqueries, federation, property paths, aggregates, and negation, and presented two non-trivial demonstrators built on top.

Two findings stand out. First, the migration to the official \ac{imdb} dataset transformed the collaboration graph from a disjoint union of one cluster per title into a densely connected component: a Bacon-number ball expanding from Kevin Bacon covers 92\% of all 38{,}067 persons within three hops, recovering the structural property that gives the demonstrator its meaning. Second, the \ac{shacl} validation surfaced a real mapping gap that the ontology alone could not catch: 25 titles whose \texttt{titleType} value (\texttt{short}, \texttt{video}, \texttt{tvMovie}, \texttt{tvSpecial}, \texttt{videoGame}) is not covered by the \ac{r2rml} mapping appear as ``phantom'' \texttt{CreativeWork} instances under \ac{rdfs} inference. The two \ac{shacl} runs --- without and with inference --- give a clean numerical illustration of the \ac{owa} / \ac{cwa} contrast: 105{,}964 violations under pure \ac{cwa} vs.\ 75 violations once subclass reasoning is applied, with the latter pointing at a specific, actionable data-quality issue.

Several follow-up directions remain open. The mapping could be extended to type the missing \texttt{titleType} values, eliminating the phantom CreativeWorks. The Bacon-number demonstrator could be packaged as a public-facing widget over the Brwsr deployment. Federation could go beyond Wikidata enrichment of birth dates towards constructing combined queries across multiple endpoints. The \ac{kg} itself, once stable, becomes a natural substrate for downstream applications such as recommendation, visualisation, or natural-language-to-\ac{sparql} interfaces.
