\newpage
\section{Introduction}

CineExplorer is an intelligent exploration platform designed to integrate and annotate autonomous information systems via an ontology. To move beyond traditional film databases, the system requires a \gls{kg} architecture that reveals the complex connections within the world of cinema. While drawing primarily on the \gls{imdb}~\cite{imdb-noncommercial} relational dataset, the system is designed to integrate and align with external data sources such as Wikidata~\cite{wikidata} to resolve ambiguities and enrich entity profiles with missing production and contextual attributes.

CineExplorer offers a dynamic interface that provides detailed information on film and actor profiles, with navigation directly controlled by the \gls{kg}. Thanks to the ontology developed, CineExplorer will offer graph-based exploration capabilities, structured around three main areas:
\begin{enumerate}
    \item CineExplorer combines data from \gls{imdb} with Wikidata to resolve semantic ambiguities. This enhancement process allows contextual and production attributes that are absent from the initial dataset to be integrated, thereby ensuring the completeness of the entity set.
    \item Beyond simple searches, users will be able to discover the link between actors based on the concept of Bacon Numbers. Using the graph, the application will calculate semantic paths to answer complex questions such as determining how many films separate two actors.
    \item The interface will be generated by the graph. Navigation interfaces will be dynamic and based on the ontology developed.
\end{enumerate}

This report covers two milestones. Milestone 2 describes the design and implementation of the ontology, including the main design decisions, the adopted OWL 2 profile, and the engineering process. Milestone 3 describes how the relational database is annotated using an R2RML mapping to produce an RDF knowledge graph conforming to the ontology.
