\section{Dataset Description}
\label{sec:dataset}

Earlier project milestones were prototyped on a third-party curated \ac{csv} sample of \ac{imdb}~\cite{northcoder-blog,northcoder-repo}. That sample was sufficient for the relational design and the initial ontology and mapping work, but a structural property of the slice prevented the Milestone~4 Bacon-number demonstrator from being exercised: no person in the curated sample appeared in more than one title, so the actor collaboration graph degenerated into disjoint single-title cliques and admitted no multi-hop expansion. To restore the connectivity the use case requires, the project migrated to \ac{imdb}'s official non-commercial release and rebuilt the database, ontology mapping, and knowledge graph against it. The earlier sample is preserved for reference in Appendix~\ref{app:old-dataset}; the rest of this section describes the official source.

\subsection{The Official IMDb Non-Commercial Dataset}

The upstream source is the \ac{imdb} non-commercial dataset~\cite{imdb-noncommercial}, published by \ac{imdb} for personal and non-commercial use under their data-use terms~\cite{imdb-terms}. It is distributed as seven gzip-compressed \ac{tsv} files refreshed daily, covering on the order of 12~million titles and 14~million persons. Two universal identifiers run through the entire release and become load-bearing in our design: \texttt{tconst} (title constant, e.g.\ \texttt{tt0073195}) and \texttt{nconst} (name constant, e.g.\ \texttt{nm0000102}). They appear as foreign keys throughout the relational design (Section~\ref{sec:infosystem}) and as the \ac{iri} templates of the knowledge graph (Section~\ref{sec:mapping}).

The seven files and their roles in the CineExplorer relational schema are summarised in Table~\ref{tab:imdb-files}.

\begin{table}[H]
\centering
\small
\begin{tabular}{p{0.20\linewidth} p{0.32\linewidth} p{0.40\linewidth}}
\hline
\textbf{File} & \textbf{Key columns} & \textbf{Role in our schema} \\
\hline
\texttt{title.basics}     & tconst, primaryTitle, titleType, startYear, runtimeMinutes, genres & Identifies titles; primary feeder for the \texttt{title} relation. \\
\texttt{title.akas}       & titleId, ordering, region, language, types, attributes             & Localised titles for \texttt{title\_aka}. \\
\texttt{title.crew}       & tconst, directors, writers                                         & Pre-aggregated director and writer lists; expanded into rows of \texttt{talent\_title}. \\
\texttt{title.episode}    & tconst, parentTconst, seasonNumber, episodeNumber                  & Series-to-episode parentage for \texttt{title\_episode}. \\
\texttt{title.principals} & tconst, ordering, nconst, category, job, characters                & One row per credited principal — source of \texttt{title\_principal} and the playable-role data. \\
\texttt{title.ratings}    & tconst, averageRating, numVotes                                    & Popularity signal; also drives the slicing criterion in Section~\ref{subsec:meaningful-examples}. \\
\texttt{name.basics}      & nconst, primaryName, birthYear, deathYear, primaryProfession, knownForTitles & The person catalogue, source of \texttt{talent}. \\
\hline
\end{tabular}
\caption{The seven \ac{tsv} files of the \ac{imdb} non-commercial release and their role in the CineExplorer relational schema. \texttt{tconst} and \texttt{nconst} are the canonical keys for titles and persons throughout.}
\label{tab:imdb-files}
\end{table}

The release ships pre-aggregated comma-separated lists in three places (\texttt{genres}, \texttt{directors}, \texttt{writers}, \texttt{knownForTitles}); our relational design normalises each of these into a junction table, restoring 1NF and recovering the join paths that the ontology mapping depends on. The full normalisation is described in Section~\ref{sec:infosystem}.

\subsection{Meaningful Examples}
\label{subsec:meaningful-examples}

The full \ac{imdb} release is too large to deploy on a laptop, and its long tail consists overwhelmingly of obscure works with very few ratings. We therefore work on a controlled slice that preserves the dense head of the distribution and is sufficient to exercise every component of the platform end-to-end. The slice serves as the running set of meaningful examples throughout the rest of this report.

\paragraph{Slice criterion.} The slice is defined by a single rule: \emph{rank all rated titles by \texttt{numVotes} in descending order, take the top $N$}. Default $N = 5{,}000$. Two properties matter for the use case:

\begin{enumerate}
\item \textbf{Popularity-driven cut-off.} \ac{imdb}'s metadata is densest for popular productions --- they accumulate more credited principals, more genre tags, and more alternative-title rows. The slice concentrates the credit mass that the actor collaboration graph depends on at the head of the distribution.
\item \textbf{Determinism.} A re-download of \texttt{title.ratings} reproduces the same top-$N$ identifiers exactly, so the slice is fully reproducible from the same dataset snapshot. The threshold is a function of the data, not a manual choice.
\end{enumerate}

The alternatives we considered --- a uniform random sample, a date-range cut, or a hand-picked seed expanded by graph traversal --- each gave up at least one of these properties. Random sampling diluted the collaboration density; a date cut over-represented one era; a seeded crawl made the slice impossible to reproduce. The numVotes-ranked top-$N$ rule is the simplest cut that preserves both. A global cut-off naturally yields a film-heavy mix (films attract more votes than individual TV episodes), but the resulting type distribution is itself informative --- the slice reflects the popularity skew that motivates the use case.

\paragraph{Slice statistics.} At $N = 5{,}000$, the resulting slice contains:

\begin{table}[H]
\centering
\small
\begin{tabular}{lrl}
\hline
\textbf{Source file} & \textbf{Rows after slice} & \textbf{Notes} \\
\hline
\texttt{title.basics}     & 5{,}000   & joined on top-$N$ tconsts \\
\texttt{title.ratings}    & 5{,}000   & ranking source for the slice \\
\texttt{title.principals} & 105{,}964 & all credit rows for retained titles \\
\texttt{title.crew}       & 5{,}000   & retained for FK closure \\
\texttt{title.episode}    & 139       & only those with parent in slice \\
\texttt{title.akas}       & 256{,}858 & only for retained titles \\
\texttt{name.basics}      & 38{,}067  & only persons referenced by retained titles \\
\hline
\end{tabular}
\caption{Slice statistics for $N=5{,}000$. Joined rows from \texttt{title.basics} and \texttt{title.ratings} populate the \texttt{title} relation; \texttt{title.principals} feeds the \texttt{principal} weak entity.}
\label{tab:slice-stats}
\end{table}

Within the 5{,}000 retained titles, the type breakdown is: \texttt{movie}~4{,}214, \texttt{tvSeries}~542, \texttt{tvEpisode}~139, \texttt{tvMiniSeries}~80, \texttt{video}~9, \texttt{short}~5, \texttt{tvMovie}~5, \texttt{tvSpecial}~3, \texttt{videoGame}~3. Films dominate, as expected from a popularity-ranked global cut-off.

\paragraph{Connectivity diagnostic.} The slice yields 13{,}521 \emph{bridge persons} (persons credited on two or more titles), well above the threshold of 100 required for a meaningful Bacon expansion. Each person averages 2.78 credits across the slice, and the resulting collaboration graph is a single large connected component covering 92\% of all persons within three hops of Kevin Bacon (Section~\ref{sec:demonstrator}).

\begin{sloppypar}\paragraph{Pipeline.} The slicing pipeline lives under \texttt{database/etl/} and consists of three Bash stages: \texttt{download.sh} (idempotent fetch and decompression of the seven \acp{tsv}), \texttt{filter\_top\_n.sh} (global \texttt{numVotes} ranking, top-$N$ tconst selection, joins through the seven \acp{tsv}, sentinel-guarded so re-running does not recompute), and \texttt{connectivity\_diagnostic.sh} (sanity checks on collaboration density before loading). The MySQL \ac{etl} itself is a single all-in-one script \texttt{database/imdb-schema.sql}, organised as six in-session phases from canonical-table \ac{ddl} through staging-table loaders to post-load FKs and indexes (Section~\ref{sec:sql}).\end{sloppypar}

\subsection{Licensing}

The \ac{imdb} data remains subject to \ac{imdb}'s non-commercial usage conditions~\cite{imdb-terms}. The slicing scripts and the relational schema we author around them are released under the project's MIT license.
