\section{Appendix: The Earlier Curated Sample}
\label{app:old-dataset}

This appendix documents the third-party curated \ac{csv} sample of \ac{imdb} used during Milestones~M1--M3, and the structural reason it was retired in favour of the official non-commercial release at Milestone~M4. It is included for completeness and to make the migration decision auditable; the body of the report (Section~\ref{sec:dataset} onward) describes only the official source actually used in the deployed knowledge graph.

\subsection*{Source}

The sample was distributed by northCoder~\cite{northcoder-blog,northcoder-repo} as a set of \ac{csv} files derived from the official \ac{imdb} non-commercial release, repackaged as a teaching dataset for relational-database exercises. It mirrored a subset of the \ac{imdb} \ac{tsv} layout, with one row per title, one row per person, and a junction file for per-credit information.

\subsection*{Scale}

The curated sample contained:
\begin{itemize}
    \item 174 titles
    \item 1{,}441 persons
    \item 586 principal credits
    \item 74 episodes
    \item 28 distinct genres
\end{itemize}

\noindent The relational design developed against this schema (entity sets, weak-entity decomposition, junction tables) carried over to the official-\ac{imdb} migration with only minor renaming, so the M1--M3 modelling work was not wasted.

\subsection*{Why It Was Retired}

The sample was excellent for early ontology and mapping work: small enough to inspect end-to-end, large enough to exercise every triple map. It was retired at Milestone~M4 because of a single structural property of the data, not the schema:

\begin{quote}
\textit{No person in the curated sample appeared in more than one title.}
\end{quote}

\noindent For the actor collaboration graph that underpins the Bacon-number demonstrator (Section~\ref{sec:demonstrator}), this property is fatal. With every person credited on at most one title, the collaboration graph degenerates into a disjoint union of single-title cliques: 1-hop and 2-hop neighbourhoods coincide, and there is no expansion to a Bacon centre or anywhere else. The headline use case the project committed to deliver could not be exercised on this data, however well-modelled.

\subsection*{Decision}

We considered three responses: (i)~hand-augment the sample with bridging credits, (ii)~graft a second sample on top, or (iii)~start over from the official \ac{imdb} non-commercial release. Option~(iii) was chosen as the only one that preserved provenance and reproducibility: the official release is canonical, daily-refreshed, and licensed for non-commercial use~\cite{imdb-noncommercial,imdb-terms}. The \ac{r2rml} mapping required only modest refactoring, and the relational schema needed minor renaming plus one new column to track \texttt{numVotes} (the slicing key introduced in Section~\ref{subsec:meaningful-examples}).

A summary of the comparison that drove the decision is given in Table~\ref{tab:dataset-comparison}.

\begin{table}[H]
\centering
\small
\begin{tabular}{p{0.26\linewidth} p{0.32\linewidth} p{0.32\linewidth}}
\hline
\textbf{Dimension} & \textbf{Curated \ac{csv} sample} & \textbf{Official \ac{imdb} non-commercial} \\
\hline
Provenance              & Third-party \ac{csv} repackage              & Direct from \ac{imdb} \\
Refresh                 & Static snapshot                        & Daily \\
Coverage                & 174 titles, 1{,}441 persons            & $\sim$12M titles, $\sim$14M persons \\
Bridge persons          & 0                                      & 13{,}521 (in the 5{,}000-title slice) \\
Bacon demonstrator      & Not realisable                         & Realisable (single connected component) \\
Reproducibility         & Static                                 & Deterministic via top-$N$ \texttt{numVotes} \\
\hline
\end{tabular}
\caption{Comparison of the earlier curated sample and the official \ac{imdb} non-commercial dataset on the dimensions that drove the Milestone~M4 migration decision.}
\label{tab:dataset-comparison}
\end{table}
